---
title: "Continuous"
---

## ToxicR for Continuous Models

Fitting continuous models are very similar in ToxicR. Here, we swap ‘continuous’ for ‘dichotomous’ in each function call. There are a few more options, which deal with the diversity of model fiting for continuous data. Let’s look at ‘single_continuous_fit()’ and ‘ma_continuous_fit()’

```{r}
library(ToxicR)
```

```{r}
mgkg <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,3,
          3,3,3,3,3,3,3,3,3,3,3,3,9,9,9,9,9,9,9,9,9,
          9,9,9,9,9,27,27,27,27,27,27,27,27,27,27,27,27,27)
count <- c(64,58,71,69,68,67,71,58,71,67,71,65,58,71,71,71,68,60,71,
           64,64,64,60,49,64,49,60,68,63,65,60,61,60,64,57,46,60,64,64,
           53,60,64,60,60,64,59,29,29,59,61,59,30,59,30,64,47,63)

data2           <- matrix(0,nrow=5,ncol=4)
colnames(data2) <- c("Dose","Resp","N","StDev")
data2[,1] <- c(0,25,50,100,200)
data2[,2] <- c(6,5.2,4.4,2.1,1.85)
data2[,3] <- c(20,20,19,20,20)
data2[,4] <- c(1.2,1.1,0.81,0.74,0.66)
```

For a continuous fit, you get some more options. Most importantly, the ‘BMD_TYPE’ variable can be set to “rel” for relative deviation, “sd” for standard deviation, “hybrid” for the Hybrid approach of Crump (1995), and “abs” for the absolute deviation definition. Each version changes what the BMR means. For the “sd”, the BMR parameter is the number of standard deviations the mean changes from background. For “rel”, one is dealing with a percent change (increase/decrease) in the mean from background. For “abs”, one specifies a fixed change from bacrkound. The Hybrid approach “hybrid” is the only BMD approach that uses the option ‘point_p’, which is a value that sould be 0<p<0.5, and this value specifies the background proability we observe an adverse response.

Two other options of importance are given. The first is the ‘distribution’ option, which specifies the distribution of response. Your options are “normal”, “normal-ncv”, and “log-normal.” When you specify normal the error term is just σ2. When it is “normal-ncv”, you let the variability of the response change with the mean σ2μ(dose)α. Finally, “log-normal” specifies the log-normal distribution with log variance of σ2.

The last option of importance is the ‘ewald’ option, which is named after Jessica Ewald who developed “Fast BMD.” All this does is calculate Wald based confidence intervals. Given it is intended for speed, and the regulatory community has not accepted it. I would always put it to FALSE, which is the default.

The final thing to note is that ToxicR accepts either summary statistics or original data. If you have original data, you should use it! In ToxicR, you have a vector of the doses and then the response. If you have summary statistics, ‘D’ is still a vector of doses; ‘Y’ is now matrix having the mean, number on test, and standard deviation as columns of that matrix.
